---
title: "soil_prop_prelim"
author: "Sara DeLaurentis"
date: "10/30/2021"
output: html_document
---

#12/07/2021 May need to change file title soon? At some point I will separate out the different components of this document and rename.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/sbaue/coding/R_TEMPRY/Itasca_project_19-21")
```

## R Markdown
Setup: Load packages and .csv

```{r }
library(tidyverse)
library(dplyr)
library(stringr)
library(viridis)
library(magrittr)
library(ggsci)

getwd()


OA_thick <- read.csv("OA_thickness_2019_21_join.csv", header=T, na.strings=c("", " ", "NA"))
class(OA_thick)

#as.tbl(OA_thick) #set as tibble. was having issues with coercing to a dataframe (using as.data.frame)

str(OA_thick)
head(OA_thick)

OA_thick[46,11] <- as.numeric(13) #depending on where the last source file was saved from, this cell may be imported as a date. The correct value should be 13 (cm). D5A R2 (2019), A_thk = 13.

#set all data types for the dataframe

OA_thick$date %>% as.Date(format = "%m/%d/%Y")
OA_thick$site %>%  as.character()
OA_thick$treatment %>%  as.character()
OA_thick$forest_type %>%  as.character()
OA_thick$inv_lvl %>% as.numeric()
OA_thick$plot %>% as.character()
OA_thick$totalO_thk %>% as.numeric()
OA_thick$Oi_thk %>% as.numeric()
OA_thick$Oe_thk %>% as.numeric()
OA_thick$Oa_thk %>% as.numeric()
OA_thick$A_thk %>% as.numeric()
OA_thick$A_type %>% as.character()
OA_thick$source %>% as.character()
OA_thick$sample_number %>% as.character()
OA_thick$bulk_density %>% as.numeric()




str(OA_thick)
view(OA_thick)

```


 
```{r }


```



## Including Plots

Checking groups of data for cleanliness

#```{r } # commented out when not needed.
C2<- OA_thick %>% filter(treatment == "C2")#view(C2)
length(C2)

C5<- OA_thick %>% filter(treatment == "C5")
view(C5)
length(C5)

D2<- OA_thick %>% filter(treatment == "D2")
view(D2)
length(D2)

D5<- OA_thick %>% filter(treatment == "D5")
view(D5)
length(D5)




## keep this code, but I'm not going into bulk density analysis yet. This will read out the bulk density values, minus any NA values in the column.
D2.bd<- OA_thick %>% 
  filter(treatment == "D2" & !is.na(bulk_density))
view(D2.bd)

D2.bd.min <- OA_thick %>% 
  filter(treatment == "D2" & !is.na(bulk_density) & horizon != "O")
view(D2.bd.min)

```
##summarize O horizon thicknesses, by site, across 2019-2020



```{r }

#O_mean_summary
O_mean_sum <- OA_thick %>% 
group_by(site) %>% 
  drop_na(totalO_thk) %>% 
  summarize(mean_O_thk = mean(mean(totalO_thk)), SD = sd(totalO_thk), n = length(site))

O_mean_sum %<>%   #Adding in SE to the dataframe  
mutate(SE = SD/sqrt(length(SD)))
view(O_mean_sum)

#write.table(O_mean_sum, file = "O_mean_sum.csv", sep = ",")


#A_mean_summary
A_mean_sum <- OA_thick %>% 
group_by(site) %>% 
  drop_na(A_thk) %>% 
  summarize(mean_A_thk = mean(mean(A_thk)), SD = sd(A_thk), n = length(site))

A_mean_sum %<>%   #Adding in SE to the dataframe  
mutate(SE = SD/sqrt(length(SD)))
view(A_mean_sum)

#write.table(A_mean_sum, file = "A_mean_sum.csv", sep = ",")

```
## Creating an O/A horizon depth plot per site. The final goal will be to plot all available paired points, but step one is to plot the means of each against each other. 

NOTE: This represents an intermediate step and will likely be heavily edited or removed because the FINAL GOAL is to plot ALL available pairs, with means and SE overlay. See next section for beginning work on the final form!

copy and paste this text into an ideas page:
Ideas for O/A pairs: O/A depths at each hole (rep), or near each hole.


Step 1: for plotting of O/A depth means: Create dataframe of O and A horizon means.

Step 2: create a dot plot, with labeled points of different colors.


```{r }
#Step 1:
O <- O_mean_sum$mean_O_thk
A <- A_mean_sum$mean_A_thk
#view(O)
#view(A)

OA_means.df <- data.frame(v1 = c("C2A","C2B","C5A","C5B","D2A","D2B","D5A","D5B"), v2 = c("C2", "C2", "C5", "C5", "D2", "D2", "D5", "D5"), v3 = O, v4 = A)
colnames(OA_means.df) = c("site", "treatment", "O","A")


view(OA_means.df)
str(OA_means.df)

write.table(OA_means.df, file = "OA_means.csv", sep = ",")

#Step 2: (this code works but the colors are gross, and the legend looks bad)
OPlot1 <- qplot(O, A, data = OA_means.df, label = site, 
      geom=c("point", "text"),
      hjust=0, vjust=0,
      color = treatment,
      xlab = "A horizon depth (cm)",
      ylab = "O horizon depth (cm)",
      main = "Plot of O horizon vs. A horizon depths (cm) by site")
print(OPlot1)





```
##Code to create an O/A depth scatterplot 

I don't like this plot because with the inclusion of C5A and C2A, possibly even C2B, it means nothing. We know the A horizon is the result of human disturbance. For publication in the main text, I would feel much more comfortable taking out C5A and C2A.

Anyways...

Step 1: Create a dataframe of the desired O/A pairs. I will need to include 2021 data when it comes in, how do I do that?
Points to include: 2019 R1-R3 O and A horizon depths, 2019 O/A depths, 2021 R1-R3 O and A horizon depths.

  -Selection criteria: "Source" == "pit" | "oad" | "oac"


```{r }

#create data frame out of desired points:

OA.graphpairs.df <- OA_thick %>% 
filter(source == "pit" | source == "oad" | source == "oac") 

view(OA.graphpairs.df)


```


Re-used from worm_plotting code.
```{r }

O_plot.FUN <- function(data, title, xlab, ylab, xcolumn, ycolumn, color_grp, l.title){
  
  Oplot <- ggplot(data, aes(x = xcolumn, y = ycolumn)) +
  geom_point(position = position_jitter(h = 0.1, w = 0.1), shape = 16, alpha = 0.8, size = 3, aes(color = (color_grp))) +
  scale_color_npg()
  #scale_color_viridis(discrete = TRUE, option = "C")

  Oplot <- Oplot + ggtitle(title) +
    labs(color = l.title) +
    theme(plot.title = element_text(size = rel(1.8), margin = margin(b = 10)))+
    theme(axis.text.x = element_text(size = rel(1.4))) +
    theme(axis.text.y = element_text(size = rel(1.4))) 
  Oplot <- Oplot +
    labs(x= xlab, y = ylab, size = rel(1.8)) +
    theme(axis.title.x = element_text(size = rel(1.5), face = "bold", margin = margin(t = 8)))+
    theme(axis.title.y = element_text(size = rel(1.5), face = "bold", margin = margin(r = 8)))+
    theme(legend.text = element_text(size = rel(1.4))) +
    theme(legend.title = element_text(size = rel(1.4)))
  
 print(Oplot) #print
}

O_plot.FUN(OA.graphpairs.df, 
"O horizon depth vs. A horizon depth by Site", 
"O horizon depth (cm)", 
"A horizon depth (cm)",
OA.graphpairs.df$totalO_thk,
OA.graphpairs.df$A_thk,
OA.graphpairs.df$site,
"Site") 

#run the function

```




```